---
title: Evaluation of COVID-19 Hospitalization Forecasters 
subtitle: Parameterized Report
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
  This notebook is a template for evaluating COVID-19 hospitalization forecast submissions from COVIDhub. After inputting a set of parameters (forecasters, COVID signals, etc), the template yields a comprehensive report on the predictions of COVID forecasters as well as their performance compared to the ground truth. The visualizations generated by the template offer an intuitive way to compare the accuracy of forecasters across all US states.
output: 
  html_document:
    code_folding: hide
params:
  forecasters:
    label: "Forecasters:"
    choices: [Covid19Sim-Simulator, CovidActNow-SEIR_CAN, COVIDhub-ensemble, COVIDhub-trained_ensemble, CU-select, Google_Harvard-CPF, GT-DeepCOVID, IHME-CurveFit, JHU_IDD-CovidSP, JHUAPL-Bucky, JHUAPL-Gecko, JHUAPL-SLPHospEns, Karlen-pypm, LANL-GrowthRate, MOBS-GLEAM_COVID, UCLA-SuEIR, UCSB-ACTS, USACE-ERDC_SEIR, USC-SI_kJalpha]
    value: [COVIDhub-ensemble, USC-SI_kJalpha, MOBS-GLEAM_COVID, Karlen-pypm, JHUAPL-SLPHospEns]
    input: select
    multiple: TRUE
  aws_scores: "covid/preliminary-buggy-reformatted-matched-scorecards-2022-01-03.RDS"
editor_options: 
  chunk_output_type: console
---

$$\\[.4in]$$
```{r common-stuff, echo=FALSE, message=FALSE, warning=FALSE}

# Package installation
knitr::opts_chunk$set(autodep = TRUE, fig.align = "center", message = FALSE, warning = FALSE, echo=params$printcode, cache = TRUE)
library(covidcast)
library(evalcast)
library(modeltools)
library(zookeeper)
library(knitr)
library(sf)
library(Cairo)
library(tidyverse)
library(viridis)
library(ggplot2)
library(plotly)
library(here)
library(aws.s3)
library(lubridate)
```

### Retrieving Forecast Data

Every week, forecasters submit their hospitalization predictions to COVID-19 ForecastHub. In this report, we rely on an AWS bucket that contains the estimates of a handful of signals (e.g., COVID death, cases, hospitalization, etc). Furthermore, the AWS server stores an array of evaluation metrics of these forecasts (e.g., Absolute Error, Weighted Interval Score, and 80% Coverage). Alternatively, the data can be retrieved from the publicly accessible [covidcast](https://cmu-delphi.github.io/delphi-epidata/api/covidcast.html) and [covideval](https://cmu-delphi.github.io/covidcast/evalcastR/) APIs.


```{r, results = 'asis', cache=TRUE}
s3a <- get_bucket("forecast-eval", region = "us-east-2")
s3b <- get_bucket("forecasting-team-data")
scores1 <- s3readRDS("score_cards_state_hospitalizations.rds", s3a, 
                     region = "us-east-2")
scores1 <- subset(
  scores1, 
  forecaster %in% union(params$forecasters, "COVIDhub-baseline"))

# The Hub always makes forecasts on Monday (dofw 2)
# We move all forecast_dates to the following Monday and shorten the ahead
wday_shift <- function(x, base_dofw = 2) (base_dofw - wday(x)) %% 7
scores1 <- scores1 %>%
  mutate(ahead = ahead - wday_shift(forecast_date),
         forecast_date = forecast_date + wday_shift(forecast_date))

# need to pass in the right forecaster here
scores <- s3readRDS(params$aws_scores, s3b) %>%
  magrittr::extract2("reformatted.buggy.matched.scorecards") %>%
  select(ahead, geo_value, forecaster, forecast_date, target_end_date,
         actual, wis, ae, cov_80, value_50, data_source, signal, 
         incidence_period) 
  # Logan's names are brutally long...
our_pred_dates <- unique(scores$forecast_date)
forecast_dates <- our_pred_dates
aheads <- unique(scores$ahead)
geo_values <- unique(scores$geo_value)
veggies <- c("asparagus", "broccoli", "chard", "daikon", 
             "escarole", "fennel", "garlic", "horseradish",
             "jicama", "kohlrabi", "lettuce", "mushroom",
             "nori", "okra", "parsnip", "radish", "squash",
             "tomato", "watercress", "baseline1", "yuca")
names_table <- tibble(forecaster = unique(scores$forecaster), veggies = veggies)
scores <- left_join(scores, names_table) %>% 
  select(-forecaster) %>%
  rename(forecaster = veggies)

results <- scores1 %>% 
  select(ahead, geo_value, forecaster, forecast_date, target_end_date,
         actual, wis, ae, cov_80, value_50, data_source, signal, 
         incidence_period) %>%
  bind_rows(scores) %>%
  filter(forecast_date %in% forecast_dates,
         ahead %in% aheads,
         geo_value %in% geo_values) 
```

**The target forecast dates are:** <br/> `r forecast_dates` 

**The template will compile data of the following forecasters:** <br/> `r params$forecasters`. 

**For this analysis, all of Logan's forecasters have been renamed:** <br/>

```{r renamed forecasters}
kableExtra::kbl(names_table) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```



$$\\[.07in]$$

### Weighted Interval Score (relative to baseline) and 80% Coverage {.tabset}

* Move all forecasters to Monday (adjust ahead backward, forecast date forward)
* Maintains the target date
* Subset to forecast dates / aheads / geo_values that Logan forecast
* We do Geometric mean relative to COVIDHub-baseline

#### WIS by Forecast Date (GeoMean)


``` {r fd-wis}
Mean <- function(x) mean(x, na.rm = TRUE)
GeoMean <- function(x, offset = 0) exp(Mean(log(x + offset)))

facets.label = str_glue("{aheads} days ahead")
names(facets.label) = aheads

subtitle = sprintf("Forecasts made over %s to %s",
                   format(min(forecast_dates), "%B %d, %Y"),
                   format(max(forecast_dates), "%B %d, %Y"))

plot_wis <-
  plot_canonical(results, 
                 x = "forecast_date", 
                 y = "wis", 
                 aggr = GeoMean,
                 grp_vars = c("forecaster","ahead"), 
                 facet_rows = "ahead", dots = FALSE,
                 base_forecaster = "COVIDhub-baseline") + 
  labs(title = subtitle, 
       x = "Forecast Dates", 
       y = "Geometric Mean WIS") +
  #geom_point(aes(forecast_date, round(wis, digits = 2), color)), alpha = 0.05) +
  facet_wrap(~ahead, nrow = 4, labeller = labeller(ahead=facets.label)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = "center"),
        legend.position = "bottom",
        legend.title = element_blank()) + 
  scale_y_log10() +
  geom_hline(yintercept = 1, size = 1.5) +
  scale_color_viridis_d() +
  guides(color = guide_legend(ncol = 2))


ggplotly(plot_wis, tooltip="text", height=800, width= 1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```

#### WIS by Forecast Date (Mean)

```{r fd-wis-m}
plot_wis <-
  plot_canonical(results, 
                 x = "forecast_date", 
                 y = "wis", 
                 aggr = Mean,
                 grp_vars = c("forecaster","ahead"), 
                 facet_rows = "ahead", dots = FALSE,
                 base_forecaster = "COVIDhub-baseline") + 
  labs(title = subtitle, 
       x = "Forecast Dates", 
       y = "Geometric Mean WIS") +
  #geom_point(aes(forecast_date, round(wis, digits = 2), color)), alpha = 0.05) +
  facet_wrap(~ahead, nrow = 4, labeller = labeller(ahead=facets.label)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = "center"),
        legend.position = "bottom",
        legend.title = element_blank()) + 
  scale_y_log10() +
  geom_hline(yintercept = 1, size = 1.5) +
  scale_color_viridis_d() +
  guides(color = guide_legend(ncol = 2))


ggplotly(plot_wis, tooltip="text", height=800, width= 1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```

#### WIS by Ahead (GeoMean)


``` {r overall-wis}
plot_wis_a <-
  plot_canonical(results, 
                 x = "ahead", 
                 y = "wis", 
                 aggr = GeoMean,
                 grp_vars = c("forecaster"), 
                 dots = TRUE,
                 base_forecaster = "COVIDhub-baseline") + 
  labs(title = subtitle, 
       x = "Days ahead", 
       y = "Geometric Mean WIS") +
  theme_bw() +
  theme(plot.title = element_text(hjust = "center"),
        legend.position = "bottom",
        legend.title = element_blank()) + 
  geom_hline(yintercept = 1, size = 1.5) +
  scale_y_log10() +
  scale_color_viridis_d() +
  guides(color = guide_legend(ncol = 2))


ggplotly(plot_wis_a, tooltip="text", height=800, width= 1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```

#### WIS by Ahead (Mean)


``` {r overall-wis-m}
plot_wis_a <-
  plot_canonical(results, 
                 x = "ahead", 
                 y = "wis", 
                 aggr = GeoMean,
                 grp_vars = c("forecaster"), 
                 dots = TRUE,
                 base_forecaster = "COVIDhub-baseline") + 
  labs(title = subtitle, 
       x = "Days ahead", 
       y = "Geometric Mean WIS") +
  theme_bw() +
  theme(plot.title = element_text(hjust = "center"),
        legend.position = "bottom",
        legend.title = element_blank()) + 
  geom_hline(yintercept = 1, size = 1.5) +
  scale_y_log10() +
  scale_color_viridis_d() +
  guides(color = guide_legend(ncol = 2))


ggplotly(plot_wis_a, tooltip="text", height=800, width= 1000) %>% 
  layout(hoverlabel = list(bgcolor = "white"))
```


#### % Coverage by Forecast Date


``` {r fd-coverage80}
plot_cov80 <-
  plot_canonical(results, 
                 x = "forecast_date", 
                 y = "cov_80", 
                 aggr = mean,
                 grp_vars = c("forecaster","ahead"), 
                 facet_rows = "ahead",
                 dots = FALSE) +
  labs(title = subtitle, x= "Forecast date", y = "Mean Coverage 80") +
  facet_wrap(~ahead, nrow = 4, labeller = labeller(ahead = facets.label)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = "center"),
        legend.position = "bottom",
        legend.title = element_blank()) + 
  scale_color_viridis_d() +
  geom_hline(yintercept = 0.8, size = 1.5) +
  guides(color = guide_legend(ncol = 2))

ggplotly(plot_cov80, tooltip="text", height=800, width=1000) 
```

#### % Coverage by Ahead

``` {r overall-coverage80}
plot_cov80_a <-
  plot_canonical(results, 
                 x = "ahead", 
                 y = "cov_80", 
                 aggr = mean,
                 grp_vars = "forecaster", 
                 dots = TRUE) +
  labs(title = subtitle, x= "Days ahead", y = "Mean Coverage 80") +
  theme_bw() +
  theme(plot.title = element_text(hjust = "center"),
        legend.position = "bottom",
        legend.title = element_blank()) + 
  scale_color_viridis_d() +
  geom_hline(yintercept = 0.8, size = 1.5) +
  guides(color = guide_legend(ncol = 2))

ggplotly(plot_cov80_a, tooltip="text", height=800, width=1000) 
```


### Maps {.tabset}

* WIS is scaled by population
* Then take Mean over all Forecast Dates and Aheads


```{r maps-processing, warning=FALSE}
library(sf)

results_intersect <- intersect_averagers(
  scores, c("forecaster"), c("forecast_date", "geo_value")) %>%
  select(c("ahead", "geo_value", "forecaster","forecast_date", "data_source", "signal","target_end_date","incidence_period","actual","wis","ae","cov_80"))

kl <- function(q, p = .8) p*log(p/q) + (1-p)*log((1-p)/(1-q))

maps <- results_intersect %>%
  group_by(geo_value, forecaster) %>%
  summarise(wis = Mean(wis),
            cov_80 = Mean(cov_80),
            kl_80 = kl(cov_80)) %>%
  left_join(animalia::state_population, by = "geo_value") %>%
  mutate(wis =  wis / population * 1e5) %>%
  pivot_longer(c("wis", "cov_80", "kl_80"), names_to = "score") %>%
  group_by(score) %>%
  mutate(time_value = Sys.Date(),
         max = max(value), min = min(value)) %>%
  group_by(forecaster, .add = TRUE) 
keys <- maps %>% group_keys()
maps <- maps %>% group_split()

levs <- levels(maps[[1]]$score)

# for county prediction, set geo_type = "county"
maps <- purrr::map(maps, 
                   ~as.covidcast_signal(
                     .x, signal = .x$score[1], 
                     data_source = .x$forecaster[1], 
                     geo_type = "state"))

maps <- purrr::map2(
  maps, keys$score,
  ~plot(.x, 
        choro_col = scales::viridis_pal()(3),
        range = switch(.y,
                       cov_80 = c(0,1), 
                       wis = c(.x$min[1], .x$max[1]),
                       kl_80 = c(0, .x$max[1]))))
```

#### Mean Weighted Interval Score


```{r map-wis, fig.width=12, fig.height=16}
cowplot::plot_grid(plotlist = maps[keys$score == "wis"], ncol = 3)
```


#### Coverage 80


```{r map-cov80, fig.width=12, fig.height=16}
cowplot::plot_grid(plotlist = maps[keys$score == "cov_80"], ncol = 3)
```


#### KL divergence 80


```{r map-kl80, fig.width=12, fig.height=16}
cowplot::plot_grid(plotlist = maps[keys$score == "kl_80"], ncol = 3)
```
